Proving linearizability using forward simulations\thanks{An extended version of this paper including the missing proofs can be found at \cite{extended}.}

### 2017-05-05 draftCAV/queue.tex
Consolidated figures.

\begin{wrapfigure}{l}{7.2cm}
\vspace{-6mm}
\includegraphics[width=7.3cm]{fig-queue12.pdf}
%
%\vspace{2mm}
%\includegraphics[width=7cm]{fig-queue2.pdf}
\vspace{-8mm}
\caption{Simulating queue histories with $AbsQ$. An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action. The representation of a dequeue operation includes a red circle that stands for a {\tt swap} returning a non-null value, which is their linearization point.}
\label{fig:queueSim}
\vspace{-6mm}
\end{wrapfigure}
Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points). The state of $AbsQ$ after each action is pictured as a graph below the action. The nodes of this graph represent enqueue operations and the edges happens-before constraints. Each node is labeled by a value (the argument of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed. For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with argument $y$. Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.

\begin{figure}[t]
  \begin{minipage}[t]{0.4\linewidth}
    \begin{lstlisting}
function present: $\mathbb{O} \to \mathbb{B}$
function pending: $\mathbb{O} \to \mathbb{B}$
function arg: $\mathbb{V} \to \mathbb{O}$
function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$

>>>>

Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points). The state of $AbsQ$ after each action is pictured as a graph below the action. The nodes of this graph represent enqueue operations and the edges happens-before constraints. Each node is labeled by a value (the argument of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed. For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with argument $y$. Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.

\begin{figure}[t]
  \begin{minipage}[b]{0.55\linewidth}
    \begin{lstlisting}
void enq(int x) {
  i = back++; items[i] = x;
}
int deq() {
  while (1) {
    range = back - 1;
    for (int i = 0; i <= range; i++) {
      x = swap(items[i],null);
      if (x != null) return x;
} } }
    \end{lstlisting}
    \caption{The Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90}.}
    \label{fig:HerlihyWing}
    \vspace{1em}

    \includegraphics[width=7cm]{fig-queue12}
    \caption{Forward simulation with $AbsQ$. Lines depict operations,
    and circles depict call, return, and linearization point actions.}
    \label{fig:queueSim}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.4\linewidth}
    \begin{lstlisting}
function present: $\mathbb{O} \to \mathbb{B}$
function pending: $\mathbb{O} \to \mathbb{B}$
function arg: $\mathbb{V} \to \mathbb{O}$
function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
